---
title: "Imputing TB logs"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---



```{r cycle_prep_imputing_logs librairies and stuff, include = FALSE, eval = TRUE}
source("Scripts/00_setup.R")
```


```{r cycle_prep_imputing_logs setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(scipen=999)
```


## Imputing logs

The absence of a TB log is not the same when other features were logged or not.

Also missing days in the middle of a logged sequence are not the same as missing days at the beginning or end of menstruation-centered cycles.


### Imputing logs: principles.

For any $i$ that is in a "missing data" position (i.e. n_logs = 0) in a sequence of **max 3 missing days in a row** and that is **not at the beginning or the end** of the cycle_m

$x_i = interpolation_i(X_L - X_N) * F_i$

where $x_i$ is the imputed value at position $i$,
$X_L$ is the TB value on the **Last** day where anything got logged (can be 1 or 0)
$X_N$ is the TB value on the **Next** day where anything got logged (can be 1 or 0)
$F_i$ is a factor that must reflect

* the uncertainty of imputing data: we actually DO NOT know what happened 

* the total number of missing day in a row: if we are missing more days, the uncertainty increases. 

* the position in the missing day sequence: we have more uncertainties in the middle than at the edges of a missing day sequence.

We have the conditions:

$ F_i < 1 $

and if $N_S1 > N_S2$ where $N_S1$ (resp $N_S2$) = length of missing days in a row in sequence 1 (resp 2), then $F_i1 < F_i2$ 

and $ F_i < F_j $ if $i$ is more on the middle than $j$


$F_i$ of the shape of an hyperbolic cosine could be a good option that respect these conditions.

[https://en.wikipedia.org/wiki/Catenary](https://en.wikipedia.org/wiki/Catenary)


### Imputing logs: implementation.

```{r cycle_prep_imputing_logs loading d_wide}

load(file = paste0(IO$tmp_data,"d_wide.Rdata"), verbose = TRUE)

```


```{r cycle_prep_imputing_logs imputing logs}

d_wide_imputed =  t(apply(d_wide[,-1], 1, impute))
d_wide_imputed_df = data.frame(cycle_id_m = d_wide$cycle_id_m, d_wide_imputed)
colnames(d_wide_imputed_df) = colnames(d_wide)
d_wide = d_wide_imputed_df

save(d_wide, file = paste0(IO$tmp_data,"d_wide_imputed.Rdata"))
rm(d_wide_imputed, d_wide_imputed_df)
```



```{r cycle_prep_imputing_logs loading cycles, cache.lazy = FALSE}
load(paste0(IO$output_data, "cycles.Rdata"), verbose = TRUE)
```


```{r cycle_prep_imputing_logs creating d}

d = reshape(d_wide, 
            idvar = "cycle_id_m", 
            varying = list(2:ncol(d_wide)), 
            new.row.names = 1:(nrow(d_wide)*(par$Df+par$D+1)),
            direction = "long")

colnames(d) = c("cycle_id_m","cycleday_m_D","tender_breasts")
d$cycleday_m_D = d$cycleday_m_D - par$D - 1
m = match(d$cycle_id_m, cycles$cycle_id)
d$user_id = cycles$user_id[m]
d$cycle_nb_m = cycles$cycle_nb[m]
d$cycle_id_m = as.character(d$cycle_id_m)
```



```{r cycle_prep_imputing_logs saving d, cache.lazy = FALSE}
save(d, file = paste0(IO$tmp_data,"d_imputed.Rdata"))
```
