---
title: "Filtering users"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---



```{r user_filt librairies and stuff, include = FALSE, eval = TRUE}
source("Scripts/00_setup.R")
```


```{r user_filt setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(scipen=999)
```


## Users filtering based on birth control

Clue's setting allow users to declare themselves as on one of these birth control (BC) methods:

```{r user_filt loading users, echo = FALSE}
load(file = paste0(IO$input_data,"users_filtered.Rdata"), verbose = TRUE)
dim(users)
colnames(users)
```


```{r user_filt possible BCs, echo = FALSE}
table(users$birth_control)
```


```{r user_filt nb of users per BC, echo = FALSE}
agg = aggregate(user_id ~ birth_control, users, uniqueN)
agg$birth_control = factor(agg$birth_control, levels = agg$birth_control[order(agg$user_id)])
ggplot(agg, aes(x = birth_control, y = user_id, fill = birth_control)) + geom_bar(stat = "identity")
```


For the purpose of this analysis, we will only keep users that have indicated `pill`, `none`, `condoms` or `did not enter`.
We will later re-classify cycles into the following BC: `pill`, `none / condoms`, `unclear` and use the only 1st two categories for the rest of the analysis.


```{r user_filt renaming the BC columns, echo = FALSE}
users$birth_control_o = users$birth_control
users$birth_control = as.character(users$birth_control)
users$birth_control[users$birth_control %in% c("none","condoms")] = "none / condoms"
table(users$birth_control)
```


### Filtering users

```{r user_filt filtering users}
j = which(users$birth_control_o %in% c("pill","none","condoms","did not enter"))
users = users[j,]

save(users, file = paste0(IO$output_data,"users.Rdata"))
```


### Filtering cycles

```{r user_filt loading cycles}
load(file = paste0(IO$input_data,"cycles_filtered.Rdata"), verbose = TRUE)
```

```{r user_filt filtering cycles}
j = which(cycles$user_id %in% users$user_id)
cycles = cycles[j,]

save(cycles, file = paste0(IO$output_data,"cycles.Rdata"))
```


### Filtering tracking tables.

As there are many reccords, the tracking table has been split into small files.
As we are filtering users, we will also make sure that the way the tracking tables are organized is suited to keep all reccords of a user in a single table.

So, we will define a `batch` number for each user and store that user's records in the file with the appropriate batch.

We want a maximum of `r par$max_batch_size` users per batch and a minimum of `r par$min_n_batches` batches.

```{r user_filt defining a batch number for each user}

n_batch = max(par$min_n_batches, ceiling(nrow(users)/par$max_batch_size))
batch_size = ceiling(nrow(users)/n_batch)

users$batch = rep(1:n_batch, each = batch_size)[1:nrow(users)]

save(users, file = paste0(IO$output_data,"users.Rdata"))
```



```{r user_filt create directory for the csv with the triaged tracking tables}

csv_batches_folder = paste0(IO$tmp_data,"tracking_csv_batches/")
dir.create(csv_batches_folder)


```


```{r user_filt triaging days and saving as csv}

tracking_folder = paste0(IO$input_data,"tracking_filtered/")
filenames = list.files(tracking_folder)

cl = makeCluster(par$n_cores)
registerDoParallel(cl)

tic()
users_original_files = foreach(filename = filenames, .combine = rbind) %dopar% 
{
  # load file
  load(paste0(tracking_folder, filename), verbose = TRUE)
  # save the colnames of the table for later
  if(filename == filenames[1]){tracking_colnames = colnames(tracking); save(tracking_colnames, file = paste0(IO$tmp_data,"tracking_colnames.Rdata"))}
  
  # filter users (to only keep those in the users table)
  j = which(tracking$user_id %in% users$user_id)
  tracking = tracking[j,]
  tracking$batch =  users$batch[match(tracking$user_id, users$user_id)]
  
  # write records for each users according to the batch they belong
  for(b in unique(tracking$batch)){
    days = tracking[which(tracking$batch == b),]
    write_csv(days, path = paste0(csv_batches_folder,"days_",b,".csv"), append = TRUE,  col_names = FALSE)
  }
  
  # keep in the users table the original file ID for each user
  original_file_id = substr(filename,25,26)
  users_original_file = data.frame(user_id = unique(tracking$user_id), original_file_id = original_file_id)
  
  return(users_original_file)
}
toc()
stopImplicitCluster()

save(users_original_files, file = paste0(IO$tmp_data, "users_original_files.Rdata"))


```

```{r user_filt saving the original file ids}

users_o_f = aggregate(original_file_id ~ user_id ,users_original_files, paste0, collapse = ",")
users$original_files = users_o_f$original_file_id[match(users$user_id, users_o_f$user_id)]

save(users, file = paste0(IO$output_data,"users.Rdata"))
file.copy(paste0(IO$output_data,"users.Rdata"), paste0(IO$tmp_data,"users_after_filtering_days.Rdata"), overwrite = TRUE)

```




