---
title: "Defining cycle ID, cycle day and menstruation-centered cycle IDs"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---



```{r data_prep_cycle_id librairies and stuff, include = FALSE, eval = TRUE}
source("Scripts/00_setup.R")
```


```{r data_prep_cycle_id setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(scipen=999)
```


## Defining cycle IDs and menstruation-centered cycles

### Defining a cycle ID in the cycle table



```{r data_prep_cycle_id load cycle table}
load(paste0(IO$output_data,"cycles.Rdata"), verbose = TRUE)
```


```{r data_prep_cycle_id defining the cycle id}
cycles$cycle_id = paste0(cycles$user_id,"_", cycles$cycle_nb)
```


```{r data_prep_cycle_id saving cycles}
save(cycles, file = paste0(IO$output_data,"cycles.Rdata"))
file.copy(paste0(IO$output_data,"cycles.Rdata"),paste0(IO$tmp_data,"cycles_after_cycle_id.Rdata"), overwrite = TRUE)

```


### Linking the days with the cycles



```{r data_prep_cycle_id linking the days with the cycles - functions}

add_cycles_to_day_table = function(filename = filename){
  cat(filename,"\n")
  load(paste0(input_days_folder, filename), verbose = TRUE)
  
  # take the part of cycles that matches with the days users
  j = which(cycles$user_id %in% unique(days$user_id))
  cycles_sub = cycles[j,]
  
  # expand cycles for each day
  cycles_sub_exp = as.data.frame(lapply(cycles_sub, rep, cycles_sub$cycle_length))
  cycles_sub_exp$cycleday = ave(rep(1,nrow(cycles_sub_exp)), cycles_sub_exp$cycle_id, FUN =cumsum)
  cycles_sub_exp$date = cycles_sub_exp$cycle_start + (cycles_sub_exp$cycleday - 1)
  cycles_sub_exp$day_id = paste0(cycles_sub_exp$user_id, "_", cycles_sub_exp$date)
  cat(filename, " : cycle_sub_exp : done\n")
  
  # match days and cycles_sub_exp
  m = match(days$day_id, cycles_sub_exp$day_id)
  days$cycle_nb = cycles_sub_exp$cycle_nb[m]
  days$cycle_id = cycles_sub_exp$cycle_id[m]
  days$cycle_length = cycles_sub_exp$cycle_length[m]
  days$cycleday = cycles_sub_exp$cycleday[m]
  
  days$cycleday_from_end = days$cycleday - days$cycle_length - 1
  
  cat(filename, " : days table has cycle_id \n")
  
  save(days, file = paste0(output_days_folder, filename))
  cat(filename, " : days table saved \n")
}

```



```{r data_prep_cycle_id linking the days with the cycles}

#input
input_days_folder = paste0(IO$tmp_data,"days_with_n_logs/")
filenames = list.files(input_days_folder)

#output
output_days_folder = paste0(IO$tmp_data,"days_with_cycle_ids/")
if(dir.exists(output_days_folder)){unlink(output_days_folder, recursive = TRUE)}
dir.create(output_days_folder)


tic()
cl = makeCluster(min(length(filenames),par$n_cores), outfile="")
registerDoParallel(cl)
foreach(filename = filenames) %dopar% {add_cycles_to_day_table(filename = filename)}
stopCluster(cl)
tac()

```


```{r data_prep_cycle_id copying days in tmp folder}
copy_days_tmp2out(output_days_folder)
```



```{r data_prep_cycle_id checking linking the days with the cycles}
filenames = list.files(output_days_folder)
load(paste0(output_days_folder, filenames[length(filenames)]), verbose = TRUE)
head(days)
summary(days)
```




## Defining menstruation-centered cycles


We consider the luteal phase to be "fixed" so that any day within D days from the next menses will belong to the next cycle.

We also define a Df which is the number of days from the follicular phase that we want to consider when we want to standardize cycles.

The column "cycleday_m" counts the days from menstruation with 0 = first day of menstruation going from -D to +x with x = cycle_length - D.

The column "cycleday_m_D" counts the days from menstruation with 0 = first day of menstruation going from -D to +X with X = min(Df,cycle_length - D).

Also, the start of the first cycle and the end of the last cycles should be removed so that cycle_id_m, cycleday_m and cycleday_m_D = NA for those (as they are "incomplete menstruation-centered cycles").



```{r data_prep_cycle_id define menstruation-centered cycle_id - parameters}

par$Df
par$D

```


```{r data_prep_cycle_id define menstruation-centered cycle_id - function}

define_menstruation_centered_cycles = function(file = "", param = par){
  cat(file,"\n")
  load(paste0(input_days_folder,file), verbose = TRUE)
  cat(file," - dim init : ",dim(days),"\n")
  
  # cycle id and cycle number
  
  days$cycle_id_m = days$cycle_id
  days$cycle_nb_m = days$cycle_nb
  
  end.j = which(days$cycleday_from_end >= -param$D)
  days$cycle_id_m[end.j] = paste0(days$user_id[end.j],"_",days$cycle_nb[end.j]+1)
  days$cycle_nb_m[end.j] = days$cycle_nb[end.j]+1
  
  rm(end.j)
  
  # cycleday_m and cycleday_m_D
  
  days$cycleday_m = days$cycleday-1
  days$cycleday_m_D = NA
  
  j.start = which((days$cycle_id_m == days$cycle_id) & (days$cycleday_m <= param$Df))
  days$cycleday_m_D[j.start] = days$cycleday[j.start]-1
  
  j.end = which(days$cycle_id_m != days$cycle_id)
  days$cycleday_m[j.end] = days$cycleday_from_end[j.end]
  days$cycleday_m_D[j.end] = days$cycleday_from_end[j.end]
  
  
  rm(j.start, j.end)
  
  # ignore what is before or after the first or last FULL cycle_id_m 
  # (i.e. we set to NA the start of the first cycle of a user 
  # and the end of the last cycle of a user)
  
  agg = aggregate(cycle_nb ~ user_id, days, range)
  agg$cycle_id_m1 = paste0(agg$user_id, "_",agg$cycle_nb[,1])
  agg$cycle_id_m2 = paste0(agg$user_id, "_",agg$cycle_nb[,2]+1)
  
  m1 = match(days$cycle_id_m, agg$cycle_id_m1)
  days$cycle_id_m[!is.na(m1)] = NA
  days$cycle_nb_m[!is.na(m1)] = NA
  days$cycleday_m[!is.na(m1)] = NA
  days$cycleday_m_D[!is.na(m1)] = NA
  
  
  m2 = match(days$cycle_id_m, agg$cycle_id_m2)
  days$cycle_id_m[!is.na(m2)] = NA
  days$cycle_nb_m[!is.na(m2)] = NA
  days$cycleday_m[!is.na(m2)] = NA
  days$cycleday_m_D[!is.na(m2)] = NA
  
  rm(agg, m1, m2)
  
  cat(file," - dim after : ",dim(days),"\n")
  save(days, file = paste0(output_days_folder,file))
  cat(file," : done\n")
}


```



```{r data_prep_cycle_id define menstruation-centered cycle_id, eval = TRUE}

#input
input_days_folder = paste0(IO$tmp_data,"days_with_cycle_ids/")
filenames = list.files(input_days_folder)

#output
output_days_folder = paste0(IO$tmp_data,"days_with_m_centered_cycle_ids/")
if(dir.exists(output_days_folder)){unlink(output_days_folder, recursive = TRUE)}
dir.create(output_days_folder)

tic()
cl = makeCluster(min(par$n_cores, length(filenames)), outfile="")
registerDoParallel(cl)
foreach(file = filenames) %dopar%{define_menstruation_centered_cycles(file = file)}
stopImplicitCluster()
tac(chunck_name = "data_prep_cycle_id define menstruation-centered cycle_id")

```


```{r data_prep_cycle_id define menstruation-centered cycle_id copy to output folder}
copy_days_tmp2out(output_days_folder)
```



```{r define checking menstruation-centered cycle_id, eval = TRUE}

filenames = list.files(output_days_folder)
load(paste0(output_days_folder, filenames[length(filenames)]), verbose = TRUE)
summary(days)
head(days)

head(days[days$cycle_nb == 2, ])
```
