---
title: "Selecting cycles around pill transitions"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---



```{r pill_trans_select librairies and stuff, include = FALSE, eval = TRUE}
source("Scripts/00_setup.R")
```


```{r pill_trans_select setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(scipen=999)
```



```{r pill_trans_select loading cycles_m and users}
load(paste0(IO$output_data,"users.Rdata"), verbose = TRUE)
load(paste0(IO$output_data,"cycles_m.Rdata"), verbose=  TRUE)
load(paste0(IO$output_data,"stretches.Rdata"), verbose = TRUE)
load(paste0(IO$output_data,"cycles.Rdata"), verbose=  TRUE)

```


## Selecting users that have a pill transition

We actually select stretches (= continuous sequence of cycles in which the BC could be reliably determined)
- with a BC transition (either on or off pill) [excluding multiple transitions]
- with at least 4 cycles before and after the transition
- with at least 8 TB in these 8 cycles



```{r pill_trans_select selecting users}

# stretches with a BC transistion
j = which(stretches$BC %in% c("on pill","off pill"))
stretch_ids = stretches$stretch_id[j]
length(stretch_ids)


# stretches with at least 4 cycles before and after the transition
agg = aggregate(cycle_nb ~ stretch_id + BC, cycles[which(cycles$stretch_id %in% stretch_ids),], lu)
agg_min = aggregate(cycle_nb ~ stretch_id, agg, min)
j = which(agg_min$cycle_nb >= 4)
stretch_ids = agg_min$stretch_id[j]
length(stretch_ids)


# adding these values to the stretches table
agg_long = cast(agg, stretch_id ~ BC, value = "cycle_nb")
colnames(agg_long) = c("stretch_id","pill","NC")
stretches$n_pill = agg_long$pill[match(stretches$stretch_id, agg$stretch_id)]
stretches$n_NC = agg_long$NC[match(stretches$stretch_id, agg$stretch_id)]
stretches$n_cycles = stretches$n_pill + stretches$n_NC



# stretches with at least 8 TB in these 8 cycles

# first we need to compute the cycle_nb at which the transition occurs
agg = aggregate(cycle_nb ~ stretch_id + BC, cycles[which(cycles$stretch_id %in% stretch_ids),], min)
agg_max = aggregate(cycle_nb ~ stretch_id, agg, max)
stretches$cycle_nb_transition = agg_max$cycle_nb[match(stretches$stretch_id, agg_max$stretch_id)]
agg_min = aggregate(cycle_nb ~ stretch_id, agg, min)
stretches$cycle_nb_start = agg_min$cycle_nb[match(stretches$stretch_id, agg_min$stretch_id)]
agg_end = aggregate(cycle_nb ~ stretch_id, cycles[which(cycles$stretch_id %in% stretch_ids),], max)
stretches$cycle_nb_end = agg_end$cycle_nb[match(stretches$stretch_id, agg_end$stretch_id)]

cycles$cycle_nb_transition = stretches$cycle_nb_transition[match(cycles$stretch_id, stretches$stretch_id)]
j = which(abs(cycles$cycle_nb - cycles$cycle_nb_transition) <= 4)
agg = aggregate(n_TB ~ stretch_id, cycles[j,], sum)
j = which(agg$n_TB >= 8)
stretch_ids = agg$stretch_id[j]

stretches$valid_transition = stretches$stretch_id %in% stretch_ids
cycles$valid_transition = cycles$stretch_id %in% stretch_ids


cycles$transition = stretches$BC[match(cycles$stretch_id, stretches$stretch_id)]
cycles$cycle_nb_from_trans = cycles$cycle_nb - cycles$cycle_nb_transition


cycles_m_valid_transition_1 = cycles$valid_transition[match(cycles_m$cycle_id_m, cycles$cycle_id)]
cycles_m_valid_transition_2 = cycles$valid_transition[match(cycles_m$cycle_id_m, paste0(cycles$user_id,"_",cycles$cycle_nb+1))]
cycles_m$valid_transition = ifelse(cycles_m_valid_transition_1 & cycles_m_valid_transition_2, "valid",ifelse(cycles_m_valid_transition_1 | cycles_m_valid_transition_2,"border","invalid"))

cycles_m$cycle_nb_m_from_trans  = cycles$cycle_nb_from_trans[match(cycles_m$cycle_id_m, cycles$cycle_id)]

save(stretches, file = paste0(IO$output_data,"stretches.Rdata"))
file.copy(from = paste0(IO$output_data,"stretches.Rdata"), to = paste0(IO$tmp_data,"stretches_with_valid_transitions.Rdata"), overwrite = TRUE)

save(cycles, file = paste0(IO$output_data,"cycles.Rdata"))
file.copy(from = paste0(IO$output_data,"cycles.Rdata"), to = paste0(IO$tmp_data,"cycles_with_valid_transitions.Rdata"), overwrite = TRUE)


save(cycles_m, file = paste0(IO$output_data,"cycles_m.Rdata"))
file.copy(from = paste0(IO$output_data,"cycles_m.Rdata"), to = paste0(IO$tmp_data,"cycles_m_with_valid_transitions.Rdata"), overwrite = TRUE)

```


## Looking at users that experiences an INcrease or DEcrease of their symptoms in these 8 cycles

```{r pill_trans_select increase vs decrease}

if(length(stretch_ids)>0){
  
  cycles$transition_cat = as.character(cut(cycles$cycle_nb_from_trans, breaks = c(-Inf,-4.5,-0.5,3.5,Inf), labels = c(NA, "before","after",NA)))
  j = which(!is.na(cycles$transition_cat))
  
  agg = aggregate(n_TB ~ stretch_id + transition_cat, cycles[j,], sum)
  agg_long = cast(agg, stretch_id ~ transition_cat, value = "n_TB")
  
  m = match(stretches$stretch_id, agg_long$stretch_id)
  stretches$before = agg_long$before[m]
  stretches$after = agg_long$after[m]
  stretches$TB_ratio = stretches$after / stretches$before
  
  change_cat_breaks = c(-Inf, 2^c( -10, -1.5,-0.5,0.5,1.5,10),Inf)
  
  stretches$TB_change_cat = cut(stretches$TB_ratio, 
                                breaks = change_cat_breaks, 
                                labels = c("complete disapearance","large descrease","small decrease","no change","small increase","large increase","complete appearance"))
  
  
  v_stretches = stretches[which(stretches$valid_transition),]
  
  table_ratio = table(round(log2(v_stretches$TB_ratio)))
  table_ratio
  
  round(100*table_ratio/sum(table_ratio),1)
  
  table_ratio_x_BC = table(v_stretches$BC, round(log2(v_stretches$TB_ratio)))
  table_ratio_x_BC
  round(100*table_ratio_x_BC/apply(table_ratio_x_BC,1,sum),1)
  
  
  table_TB_change_cat_x_BC = table(v_stretches$BC, v_stretches$TB_change_cat)
  table_TB_change_cat_x_BC
  round(100*table_TB_change_cat_x_BC/apply(table_TB_change_cat_x_BC,1,sum),1)
  
  
  table_TB_change_cat_x_BC_df = melt(table_TB_change_cat_x_BC)
  colnames(table_TB_change_cat_x_BC_df) = c("BC","TB_change_cat","n_users")
  table_TB_change_cat_x_BC_df = table_TB_change_cat_x_BC_df[table_TB_change_cat_x_BC_df$BC %in% c("on pill", "off pill"),]
  
  table_TB_change_cat_x_BC_df$TB_change_cat_wrap = str_wrap(table_TB_change_cat_x_BC_df$TB_change_cat, width = 12)
  table_TB_change_cat_x_BC_df$TB_change_cat_wrap = factor(table_TB_change_cat_x_BC_df$TB_change_cat_wrap, levels = str_wrap(levels(stretches$TB_change_cat), width = 12))
  
  
  g = ggplot(table_TB_change_cat_x_BC_df, aes(x = TB_change_cat_wrap, y = n_users, fill = BC))+
    geom_bar(stat = "identity")+
    facet_grid(BC ~., scale = "free")+
    xlab("")+ylab("# of users")
  g
  
  bw = 0.5
  
  v_stretches$TB_ratio_log2 = log2(v_stretches$TB_ratio)
  value_inf = round(max(abs(v_stretches$TB_ratio_log2[!is.infinite(v_stretches$TB_ratio_log2)]), na.rm = TRUE),1)+bw/2
  v_stretches$TB_ratio_log2[is.infinite(v_stretches$TB_ratio_log2)] = value_inf*sign(v_stretches$TB_ratio_log2[is.infinite(v_stretches$TB_ratio_log2)])
  
  g_hist = ggplot(v_stretches, aes(x = TB_ratio_log2, fill = BC))+
    geom_histogram(binwidth = bw)+
    facet_grid(BC ~., scale = "free")+
    xlab("")+ylab("# of users")+
    geom_vline(xintercept = c(-value_inf+0.5,-1.5,-0.5,0.5,1.5,value_inf-0.5), col = "gray20", linetype = 2, size = 0.2)+
    geom_vline(xintercept = 0, col = "gray20", linetype = 1, size = 0.5)
  g_hist
  
  
  v_stretches$TB_ratio_log2_cut = as.numeric(as.character(cut(v_stretches$TB_ratio_log2, breaks = seq(-value_inf-bw/2,value_inf+bw/2,by = bw), labels =  seq(-value_inf,value_inf,by = bw))))
  
  hist_log2_TB_ratio_x_BC = aggregate(stretch_id ~ TB_ratio_log2_cut + BC, v_stretches, length)
  colnames(hist_log2_TB_ratio_x_BC) = c("log2_TB_ratio","transition","n_users")
  
  save(table_TB_change_cat_x_BC_df, file = paste0(IO$out_Rdata,"table_TB_change_cat_x_BC_df.Rdata"))
  save(hist_log2_TB_ratio_x_BC, file = paste0(IO$out_Rdata,"hist_log2_TB_ratio_x_BC.Rdata"))
  
  
  
  save(stretches, file = paste0(IO$output_data,"stretches.Rdata"))
  file.copy(from = paste0(IO$output_data,"stretches.Rdata"), to = paste0(IO$tmp_data,"stretches_with_valid_transitions.Rdata"), overwrite = TRUE)
  
}

```






```{r pill_trans_select selecting days from these cycles}
if(length(stretch_ids)>0){
  
  cycle_ids = cycles$cycle_id[which((!is.na(cycles$transition_cat))&(cycles$valid_transition))]
  
  input_folder = paste0(IO$output_data,"days/")
  files = list.files(input_folder)
  
  
  cl = makeCluster(par$n_cores)
  registerDoParallel(cl)
  
  days_trans = foreach(file = files, .combine = rbind) %dopar%{
    load(paste0(input_folder, file), verbose=  TRUE)
    days = days[days$cycle_id %in% cycle_ids,]
    return(days)
  }
  
  stopCluster(cl)
  
  days = days_trans
  
  m = match(days$cycle_id, cycles$cycle_id)
  days$stretch_id = cycles$stretch_id[m]
  days$cycle_nb_from_trans = cycles$cycle_nb_from_trans[m]
  days$transition = cycles$transition[m]
  days$transition_cat = cycles$transition_cat[m]
  
  m = match(days$stretch_id, stretches$stretch_id)
  days$TB_change_cat = stretches$TB_change_cat[m]
  
  
  m = match(days$cycle_id_m, cycles_m$cycle_id_m)
  days$cycle_nb_m_from_trans = cycles_m$cycle_nb_m_from_trans[m]
  
  
  save(days, file = paste0(IO$tmp_data, "days_pill_trans.Rdata"))
}

```



## Pre- and post-transition profiles


```{r pill_trans_select TB profiles at pill transitions}
if(length(stretch_ids)>0){
  
  days$TB_change_cat_simple = case_when(
    as.numeric(days$TB_change_cat) %in% c(4) ~ as.character(days$TB_change_cat),
    as.numeric(days$TB_change_cat) %in% c(1,2,3) ~ "decrease",
    as.numeric(days$TB_change_cat) %in% c(5,6,7) ~ "increase",
    TRUE ~ "undefined"
  )
  
  
  #ifelse(as.numeric(days$TB_change_cat) == 4, "no change",ifelse(as.numeric(days$TB_change_cat) < 4, "decrease","increase"))
  
  j = which((days$type == TB) & (days$TB_change_cat_simple != "undefined"))
  
  days$n_tender_breasts = 1
  agg = aggregate(n_tender_breasts ~ cycle_nb_m_from_trans + cycleday_m_D + transition + BC + TB_change_cat_simple,days[j,], sum )
  
  i = which(days$TB_change_cat_simple != "undefined")
  agg_tot = aggregate(cycle_id ~ cycle_nb_m_from_trans + transition + BC + TB_change_cat_simple ,days[i,], lu )
  agg_tot_max = aggregate(cycle_id ~  transition + TB_change_cat_simple ,agg_tot, max)
  colnames(agg_tot_max) = c("transition","TB_change_cat_simple","n_users")
  agg = merge(agg, agg_tot_max, all = TRUE)
  agg$fraction_cycles_with_TB = agg$n_tender_breasts / agg$n_users
  
  transition_profiles = agg
  
  g = ggplot(transition_profiles, aes(x = cycleday_m_D, y = fraction_cycles_with_TB, col = BC)) + geom_line()+ facet_grid(transition ~ cycle_nb_m_from_trans) 
  print(g)
  
  save(transition_profiles, file = paste0(IO$out_Rdata,"pill_transition_profiles.Rdata"))
}
```






