---
title: "Phenotyping users"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---

```{r user_phenotyping librairies and stuff, include = FALSE, eval = TRUE, cache = FALSE}
source("Scripts/00_setup.R")
```


```{r user_phenotyping knitr setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Users phenotyping


## Multidimensional scaling and clustering

We want to represent users in a lower dimensional space and potentially identify clusters based on the average profiles of users that were found to track and report symptoms consistently. 

We will do this by BC group and merge the "regular tracking" with the "pre-menstrual tracking" groups.

To compare two user's average profiles, we will use the [weighted jaccard distance](https://en.wikipedia.org/wiki/Jaccard_index). 


```{r user_phenotyping load data}
load(paste0(IO$output_data, "users.Rdata"), verbose = TRUE)
load(paste0(IO$tmp_data, "avg_profile_per_user.Rdata"), verbose = TRUE)
```




```{r user_phenotyping formating data}
c_wide = cast(avg_profile_per_user[,c("user_id","cycleday_m_D","avg_tender_breast")], user_id ~ cycleday_m_D)
c_wide[is.na(c_wide)] = 0
save(c_wide, file = paste0(IO$tmp_data, "c_wide.Rdata"))
```



```{r user_phenotyping selecting users}

j = (users$BC %in% par$BC_dict$name) &
  (users$consistency >= 0.5) &
  ((users$f_pre.menstrual_tracking >= 0.75) | (users$f_regular_tracking >= 0.75))

user_ids = users$user_id[which(j)]

c_wide = c_wide[which(c_wide$user_id %in% user_ids),]
save(c_wide, file = paste0(IO$tmp_data, "c_wide_selected_users.Rdata"))


```


```{r user_phenotyping c_wide_BC}
c_wide_BC = users$BC[match(c_wide$user_id, users$user_id)]
```



```{r user_phenotyping characteristics}
#load(file = paste0(IO$tmp_data, "c_wide_selected_users.Rdata"), verbose = TRUE)
# load(file = paste0(IO$output_data,"users.Rdata"), verbose = TRUE)


#BC = as.character(par$BC_dict$name)[1] # NC
#BC = as.character(par$BC_dict$name)[2] # pill


characteristics_both = data.frame()

for(BC in as.character(par$BC_dict$name)){
  
  k = which(c_wide_BC == BC)
  this_c_wide = c_wide[k,]
  M = this_c_wide[,-1]
  
  if(BC != "pill"){bc = "NC"}else{bc = BC}
  save(M, file = paste0(IO$tmp_data,"user_phenotyping_M_",bc,".Rdata"))
  
  
  characteristics = data.frame(user_id = this_c_wide$user_id)
  m = match(characteristics$user_id, users$user_id)
  characteristics$BC = users$BC[m]
  characteristics$age = users$age[m]
  characteristics$bmi_cat = users$bmi_cat[m]
  characteristics$bmi_cat_num = as.numeric(characteristics$bmi_cat)
  
  
  characteristics$sum = apply(M, 1, sum)
  characteristics$max = apply(M, 1, max)
  characteristics$time = apply(M, 1, function(x){weighted.mean(x = par$cycleday_m_D, w = x)})
  characteristics$time_max = apply(M, 1, function(x){mean(par$cycleday_m_D[x==max(x)])})
  characteristics$n_days = apply(M, 1, function(x){sum(x>0)})
  characteristics$time_first_symptoms = apply(M, 1, function(x){min(c(-par$D:par$Df)[x>0], na.rm = TRUE)})
  characteristics$time_last_symptoms = apply(M, 1, function(x){max(c(-par$D:par$Df)[x>0], na.rm = TRUE)})
  characteristics$discontinuity = characteristics$time_last_symptoms -  characteristics$time_first_symptoms  - characteristics$n_days 
  characteristics$variability = apply(M, 1, function(x){sum(abs(diff(x)))}) # sum(rle(sign(diff(x)))$values == -1)
  
  save(characteristics, file = paste0(IO$tmp_data,"user_phenotyping_characteristics_",bc,".Rdata"))
  
  characteristics_both = rbind(characteristics_both, characteristics)
  
  print(ggplot(characteristics, aes(x = age, y = time_max)) + geom_point(alpha = 0.05) + geom_smooth()+ggtitle(BC))
  print(ggplot(characteristics, aes(x = age, y = time_last_symptoms)) + geom_point(alpha = 0.01)+ geom_smooth()+ggtitle(BC))
  print(ggplot(characteristics, aes(x = age, y = n_days)) + geom_point(alpha = 0.01)+ geom_smooth()+ggtitle(BC))
  
} 


characteristics_both_long = melt(characteristics_both, id.vars = c("user_id","BC","age","bmi_cat","bmi_cat_num"))
ggplot(characteristics_both_long, aes(x = value, col = BC, fill = BC))+
  geom_histogram(aes(y = ..density..),position = "identity", alpha = 0.4, bins = 26)+
  facet_wrap(variable ~., scale = "free")

ggplot(characteristics_both, aes(x = n_days, fill = BC))+
  geom_histogram(binwidth = 1)+
  facet_grid(BC ~., scale = "free")


ggplot(users[users$user_id %in% characteristics_both$user_id,], aes(x = consistency, fill = BC))+
  geom_histogram(bins = 10)+
  facet_grid(BC ~., scale = "free")

```



```{r user_phenotyping adding characteristics to users table}

col_to_add = colnames(characteristics_both)
j = which(col_to_add %in% colnames(users))
if(length(j)>0){col_to_add = col_to_add[-j]}
m = match(users$user_id, characteristics_both$user_id)
for(col in col_to_add){
  eval(parse(text = paste0("users$",col," = characteristics_both$",col,"[m]")))
}

save(users, file = paste0(IO$output_data,"users.Rdata"))
file.copy(from = paste0(IO$output_data,"users.Rdata"), to = paste0(IO$tmp_data,"users_with_characteristics.Rdata") )


```


### Distance (weighted Jaccard)

```{r user_phenotyping distance}

source("Scripts/00_functions_distance.R")


for(BC in as.character(par$BC_dict$name)){
  
  if(BC != "pill"){bc = "NC"}else{bc = BC}
  load(file = paste0(IO$tmp_data,"user_phenotyping_M_",bc,".Rdata"), verbose = TRUE)
  
  # DISTANCE
  dist = user_distance(M = M)
  dist = as.dist(dist)
  
  save(dist, file = paste0(IO$tmp_data,"user_phenotyping_dist_",bc,".Rdata"))
  
}  

```

### Multidimensional scaling


```{r user_phenotyping multidimensional scaling}


for(BC in as.character(par$BC_dict$name)){
  
  if(BC != "pill"){bc = "NC"}else{bc = BC} 
  load(file = paste0(IO$tmp_data,"user_phenotyping_dist_",bc,".Rdata"), verbose = TRUE)
  load(file = paste0(IO$tmp_data,"user_phenotyping_characteristics_",bc,".Rdata"), verbose = TRUE)
  
  
  # MULTIDIMENSIONAL SCALING
  
  mds = cmdscale(dist, k = 26, eig = TRUE)
  #screeplot
  barplot(mds$eig[1:10])
  eig_df = data.frame(x = 1:length(mds$eig), eig = mds$eig)
  # data frame for viz
  mds_df = data.frame(characteristics, mds$points)
  
  print(ggplot(mds_df, aes(x = X1, y = X2)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = sum)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = sum)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = max)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = max)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = time)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = time)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = time_max)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = time_max)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = n_days)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = n_days)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = time_first_symptoms)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = time_first_symptoms)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = time_last_symptoms)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = time_last_symptoms)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = discontinuity)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = discontinuity)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = variability)) + geom_point(alpha = 0.5)+ggtitle(BC))
  print(ggplot(mds_df, aes(x = X1, y = X3, col = variability)) + geom_point(alpha = 0.5)+ggtitle(BC))
  
  
  k = which(round(mds$points[,1],1) == 0) 
  o = order(mds$points[k,2])
  
  first = k[o[1:5]]
  last = k[o[(length(k)-5):length(k)]]
  
  ggplot(mds_df, aes(x = X1, y = X2, col = time)) + geom_point(alpha = 0.5)+
    geom_point(data = mds_df[first,], aes(x = X1, y = X2), col = "deeppink", shape = "+", size = 3)+
    geom_point(data = mds_df[last,], aes(x = X1, y = X2), col = "orange", shape = "+", size = 3)
  
  
  matplot(par$cycleday_m_D, t(M[first,]), type = "l", lty = 1, col = "deeppink")
  matplot(par$cycleday_m_D, t(M[last,]), type = "l", lty = 2, col = "orange", add = TRUE)
  
  par$cycleday_m_D * M[first[1],]
  mean(as.numeric(par$cycleday_m_D * M[first[1],]))
  
  mds_df$time[first]
  mds_df$time[last]
  
  j = which(this_c_wide$user_id == special_pill_user)
  as.matrix(dist)[j,]
  print(ggplot(mds_df, aes(x = X1, y = X2, col = time)) + geom_point(alpha = 0.5)+
          geom_point(data = mds_df[j,], aes(x = X1, y = X2), col = "deeppink", shape = "+", size = 3))
  
  
  print(ggplot(mds_df, aes(x = X1, y = X2, col = variability)) + geom_point(alpha = 0.5))
  print(ggplot(mds_df, aes(x = X1, y = X2, col = variability)) + geom_point(alpha = 0.5)+
          geom_point(data = mds_df[j,], aes(x = X1, y = X2), col = "deeppink", shape = "+", size = 3))
  
  # interactive 3d views
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3,
          color = ~X4,
          type = "scatter3d", mode = "markers",
          marker = list(size = 4), title = )%>% layout(title = BC)
  
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~n_days, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~sum, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~max, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~time, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~time_max, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~time_first_symptoms, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~time_last_symptoms, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~discontinuity, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~variability, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~age, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  plot_ly(mds_df, x = ~X1, y = ~X2, z = ~X3, 
          color = ~bmi_cat_num, type = "scatter3d", mode = "markers",
          marker = list(size = 4))
  
  
  
  
  # t-SNE
  
  rtsne_out = Rtsne(dist, is_distance = TRUE, perplexity = 30, theta = 0.5)
  rtsne_out_df = data.frame(characteristics, rtsne_out$Y)
  
  print(ggplot(rtsne_out_df, aes(x = X1, y = X2, col = sum)) + geom_point(alpha = 0.5))
  
  print(ggplot(rtsne_out_df, aes(x = X1, y = X2, col = max)) + geom_point(alpha = 0.5))
  
  print(ggplot(rtsne_out_df, aes(x = X1, y = X2, col = time)) + geom_point(alpha = 0.5))
  
  print(ggplot(rtsne_out_df, aes(x = X1, y = X2, col = n_days)) + geom_point(alpha = 0.5))
  
  print(ggplot(rtsne_out_df, aes(x = X1, y = X2, col = time_max)) + geom_point(alpha = 0.5))
  
  
  rtsne_out_df$color = rgb(
    red = re_scale_range(-rtsne_out_df$time_max), 
    green = re_scale_range(rtsne_out_df$n_days), 
    blue = re_scale_range(rtsne_out_df$time),
    maxColorValue = 1)
  
  plot(rtsne_out_df$X1, rtsne_out_df$X2,  col = rtsne_out_df$color, pch = 16, cex = 0.5)
  
  
  if(BC != "pill"){bc = "NC"}else{bc = BC}
  save(mds_df, eig_df,rtsne_out_df, file = paste0(IO$tmp_data,"user_phenotyping_mds_",bc,".Rdata"))
}


```

### Clustering

```{r user_phenotyping clustering}

for(BC in as.character(par$BC_dict$name)){
  
  if(BC != "pill"){bc = "NC"}else{bc = BC} 
  load(file = paste0(IO$tmp_data,"user_phenotyping_dist_",bc,".Rdata"), verbose = TRUE)
  load(file = paste0(IO$tmp_data,"user_phenotyping_characteristics_",bc,".Rdata"), verbose = TRUE)
  
  
  # CLUSTERING
  
  ## optimal number of clusters
  silh_k = data.frame()
  for(k in 2:10){
    pam_k = pam(x = dist, k = k, diss = TRUE)
    silh_df = data.frame(pam_k$silinfo$widths, x =  nrow(pam_k$silinfo$widths):1)
    g_silh = ggplot(silh_df,aes(x = x, y = sil_width, fill = factor(cluster)))+geom_bar(stat = "identity")+coord_flip()+guides(col = FALSE)+ggtitle(BC)
    medoid_df = avg_profile_per_user[which(avg_profile_per_user$user_id %in% characteristics$user_id[pam_k$medoids]),]
    g_med = ggplot(medoid_df, aes(x = cycleday_m_D, y = avg_tender_breast, col = user_id))+geom_line()+guides(col = FALSE)+facet_grid(user_id~.)+ggtitle(BC)
    grid.arrange(g_med, g_silh, nrow = 1)
    silh_k = rbind(silh_k, data.frame(n_clust = k, avg_width = pam_k$silinfo$avg.width))
  }

  print(ggplot(silh_k, aes(x =n_clust, y = avg_width))+ 
    geom_point()+geom_line()+ggtitle(BC)+ylim(c(0,1)))
  
}

```

We observe that the average silhouette width is still low even at its maximum > users cannot be separated very well.

