---
title: "Cycle clustering with condition on n_logs"
author: "Laura Symul"
date: "12/10/2018"
output: html_document
---




##  CYCLES CLUSTERING ENOUGH LOGS

Clustering at the cycle level filtering for cycles that have at least 3 days with logs in [-7,-1] and in [1,7]
 

### Selecting cycles with at least 1 logged tender breasts and with enough days logged

```{r cycle_clustering_enough_logs selecting cycles with at least 1 logged TB and enough days logged}

days_filenames = list.files(paste0(IO$Rdata, "days/"))

registerDoParallel(par$n_cores)

tic()
days = foreach(days_filename = days_filenames, .combine = rbind)%dopar%{
  load(paste0(IO$Rdata,"days/",days_filename), verbose = TRUE)
  
  #enougn logs pre-menstrual phase
  agg = aggregate(cycleday_m ~ cycle_id_m, days[(days$type == "n_logs") & (days$cycleday_m %in% -7:-1),],lu )
  cycles_ids = agg$cycle_id_m[agg$cycleday_m>=3]
  #enougn logs post-menstrual phase
  agg = aggregate(cycleday_m ~ cycle_id_m, days[(days$type == "n_logs") & (days$cycleday_m %in% 1:7),],lu )
  cycles_ids = intersect(cycles_ids, agg$cycle_id_m[agg$cycleday_m>=3])
  
  row_keep = (days$type == TB) & (days$cycle_id_m %in% cycles_ids)
  col_keep = c("type","user_id",
               "cycle_nb","cycle_id","cycleday",
               "cycle_nb_m","cycle_id_m","cycleday_m","cycleday_m_D",
               "cycle_nb_mb","cycle_id_mb","cycleday_mb","cycleday_mb_D")
  days = days[row_keep,col_keep]
  return(days)
}
toc()
stopImplicitCluster()

save(days, file = paste0(IO$Rdata,"days_TB_only_cycles_with_enough_logs.Rdata"))
```

```{r cycle_clustering_enough_logs load days_TB_only_cycles_with_enough_logs, eval = FALSE}
load(file = paste0(IO$Rdata,"days_TB_only_cycles_with_enough_logs.Rdata"))
```





```{r cycle_clustering_enough_logs checking the overall distribution of TB logs}
agg  = aggregate(type ~ cycleday_m_D, days, length)
ggplot(agg, aes(x = cycleday_m_D, y = type)) + geom_line()
```

Clearly, the peak on Day 0 (first day of menstruation) has *substantially decreased* when considering cycles in which users have logged at least 3 days in the 7 days prior to menstruation.




### Formatting days for clustering

```{r cycle_clustering_enough_logs formatting days for clustering}
k = which(!is.na(days$cycleday_m_D))
d = days[k,]
d$tender_breasts = 1

d_wide = reshape(d[,c("cycle_id_m","cycleday_m_D","tender_breasts")],
                        idvar = "cycle_id_m",timevar = "cycleday_m_D",
                        direction = "wide")
d_wide[is.na(d_wide)] = 0

# re-order the columns by cycleday_m_D
o = order(as.numeric(sapply(strsplit(colnames(d_wide)[2:ncol(d_wide)],"\\."), "[[", 2)))
d_wide = data.frame(cycle_id_m = d_wide$cycle_id_m, d_wide[,o+1] )
rm(o, k)

head(d_wide)


head_d_wide = d_wide[1:20,]
colnames(head_d_wide) = gsub("tender_breasts..","-",colnames(head_d_wide), fixed = TRUE)
colnames(head_d_wide) = gsub("tender_breasts.","",colnames(head_d_wide), fixed = TRUE)
head_d_wide

```

```{r cycle_clustering_enough_logs example of cycle raw data for presentation, eval = FALSE}
write.table(head_d_wide, file = "data/example_day_values_for_cycle_clustering_enough_logs.txt", col.names = TRUE, row.names = FALSE, sep = "\t",quote = FALSE)
```




### Ideal number of clusters ?

```{r cycle_clustering_enough_logs cycle clusters - exploring number of clusters}

cycle_ids = unique(d$cycle_id_m)
WSS_k_tot.withinss = c()
n_centers = 2:7
tic()
for(n_center in n_centers){
  cat(n_center,"\n")
  kmeans_type = kmeans(d_wide[,-1], centers = n_center, nstart = 20)
  if(n_center == 2){WSS_k_tot.withinss = c(kmeans_type$totss, kmeans_type$tot.withinss) }
  else{ WSS_k_tot.withinss = c(WSS_k_tot.withinss, kmeans_type$tot.withinss )}
  
  d$kmean_group = kmeans_type$cluster[match(d$cycle_id_m, d_wide$cycle_id_m)]
  
  d = d[order(d$kmean_group),]
  d$cycle_id_m_factors_kmeans = factor(d$cycle_id_m, levels = unique(d$cycle_id_m))
  
  #### plots
  # first showing some examples
  set.seed(6)
  d_subset = d[d$cycle_id_m %in% cycle_ids[sample(1:length(cycle_ids),50)],]
  # black dots
  print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m)) + geom_point() )
  # colored dots
  print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m, col = factor(kmean_group))) + geom_point()+ guides(col = FALSE))
  # ordered colored dots
  d_subset = d_subset[order(d_subset$kmean_group),]
  d_subset$cycle_id_m = factor(d_subset$cycle_id_m, levels = unique(d_subset$cycle_id_m))
  print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m, col = factor(kmean_group))) + geom_point()+ guides(col = FALSE))
  
  # then showing the profiles for the clusters
  agg = aggregate(tender_breasts ~ cycleday_m_D + kmean_group, d, sum)
  print(ggplot(agg, aes(x = cycleday_m_D, y =tender_breasts, col = factor(kmean_group))) + geom_line() + facet_grid(kmean_group ~.))
}
toc()
df = data.frame(n_clust = c(1,n_centers), WSS = WSS_k_tot.withinss)
ggplot(df, aes(x =n_clust, y = WSS)) + geom_bar(stat = "identity", fill = "steelblue2")

rm(df,cycle_ids, kmeans_type, d_subset, agg, n_center, n_centers, WSS_k_tot.withinss)
```





There is not a very clear "elbow", but if any, it is at number of cluster = 2. 
We see that indeed, the main 2 clusters are

* Menstrual tender-breasts (sharp increase on Day1)
* Pre-menstrual tender-breasts (soft curve before menstruation)

One could even consider a third group (that appears in the n_clust = 4) with a soft curve within menstruations...

But for now, let's consider 2 clusters.

### 2 clusters

```{r cycle_clustering_enough_logs cycle clusters - 2 clusters}
n_center = 2
kmeans_type = kmeans(d_wide[,-1], centers = n_center, nstart = 20)

d$kmean_group_num = kmeans_type$cluster[match(d$cycle_id_m, d_wide$cycle_id_m)]

agg_clust = aggregate(tender_breasts ~ cycleday_m_D + kmean_group_num, d, sum)
# percentage of reported symptoms
agg =  aggregate(cycle_id ~ kmean_group_num, d, lu)
colnames(agg) = c("kmean_group_num","n_cycles")
agg_clust = merge(agg_clust, agg, all = TRUE)
agg_clust$perc_tender_breasts = agg_clust$tender_breasts/agg_clust$n_cycles
# renaming the clusters as "menstrual" and "pre-menstrual"
agg = aggregate(perc_tender_breasts ~ kmean_group_num, agg_clust, max)
agg_clust$max_perc_tender_breasts = agg$perc_tender_breasts[match(agg_clust$kmean_group_num, agg$kmean_group_num)]
agg = agg_clust[agg_clust$perc_tender_breasts == agg_clust$max_perc_tender_breasts,]
agg$kmean_group = c("pre-menstrual","menstrual")[(agg$cycleday_m_D == 0)+1]
agg$kmean_group = factor(agg$kmean_group, levels = c("menstrual","pre-menstrual"))
agg_clust$kmean_group = agg$kmean_group[match(agg_clust$kmean_group_num, agg$kmean_group_num)]

d$kmean_group = agg$kmean_group[match(d$kmean_group_num, agg$kmean_group_num)]
d = d[order(d$kmean_group_num),]
d$cycle_id_m_factors_kmeans = factor(d$cycle_id_m, levels = unique(d$cycle_id_m))

#### plots
# first showing some examples
set.seed(6)
cycle_ids = unique(d$cycle_id_m)
d_subset = d[d$cycle_id_m %in% cycle_ids[sample(1:length(cycle_ids),50)],]
# black dots
print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m)) + geom_point() )
# colored dots
print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m, col = kmean_group)) + geom_point()+ guides(col = FALSE)+ scale_color_manual(values = cols$clusters))
# ordered colored dots
d_subset = d_subset[order(d_subset$kmean_group),]
d_subset$cycle_id_m = factor(d_subset$cycle_id_m, levels = unique(d_subset$cycle_id_m))
print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_id_m, col = kmean_group)) + geom_point()+ guides(col = FALSE)+ scale_color_manual(values = cols$clusters))

# then showing the profiles for the clusters
print(ggplot(agg_clust, aes(x = cycleday_m_D, y =tender_breasts, col = kmean_group)) + geom_line() + facet_grid(kmean_group ~.)+ scale_color_manual(values = cols$clusters))
print(ggplot(agg_clust, aes(x = cycleday_m_D, y =perc_tender_breasts, col = kmean_group)) + geom_line() + facet_grid(kmean_group ~.)+ scale_color_manual(values = cols$clusters))



rm(cycle_ids, d_subset, agg, n_center)
```




```{r cycle_clustering_enough_logs save cycle clustering}
cycle_clusters = data.frame(cycle_id_m = d_wide$cycle_id_m, kmean_group_num = kmeans_type$cluster)
cycle_clusters$kmean_group = unique(agg_clust[order(agg_clust$kmean_group_num),c("kmean_group_num","kmean_group")])[cycle_clusters$kmean_group_num,2]

save(cycle_clusters, file = paste0(IO$Rdata, "cycles_with_clusters_enough_logs.Rdata"))

rm(kmeans_type)

```


```{r cycle_clustering_enough_logs cycle clusters - figure 2A}

g = ggplot(agg_clust, aes(x = cycleday_m_D, y =perc_tender_breasts, col = kmean_group))+ 
  geom_vline(xintercept = 0, col = "gray90", size = 1.5) + 
  geom_point(size = 0.6)+geom_line() +
  scale_y_continuous(labels = scales::percent)+ scale_x_continuous(breaks = par$x.axis)+ 
  scale_color_manual(values = cols$clusters)+
  xlab("cycleday from 1st day of menstruation") + ylab("% of reported tender breasts")+
  guides(color = FALSE)

g

ggsave(filename = paste0(IO$panels, "2A_cycle_clusters.pdf"), plot = g, width = viz$full_width/9*2, height = viz$full_width/5)


agg = aggregate(n_cycles ~ kmean_group, agg_clust, min)
agg$perc_cycles = agg$n_cycles/sum(agg$n_cycles)
agg$kmean_group_short = c("M","Pre-M")[(agg$kmean_group == "pre-menstrual")+1]
g = ggplot(agg, aes(x = kmean_group_short, y =perc_cycles, fill = kmean_group)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent, position = "right")+ 
  scale_fill_manual(values = cols$clusters)+
  xlab("") + ylab("")+ ggtitle("% cycles")+
  guides(fill = FALSE)+ theme(plot.title = element_text(size=10))
g
ggsave(filename = paste0(IO$panels, "2A_n_cycles_in_clusters.pdf"), plot = g, width = viz$full_width/9, height = viz$full_width/5)


rm(g,  agg, agg_clust)
```


```{r cycle_clustering_enough_logs cycle clusters examples, eval = FALSE}

# selecting 3 users: 1 only M, 1 only pre-M, 1 mixed
agg = aggregate(kmean_group_num ~ user_id, d, lu)
colnames(agg) = c("user_id","n_cycle_clusters")
users = merge(users, agg, all = TRUE)

agg = aggregate(kmean_group_num ~ user_id + cycle_id_m + kmean_group, d, min)
agg$kmean_group_short = c("M","PreM")[(agg$kmean_group == "pre-menstrual")+1]
agg = aggregate(cycle_id_m ~ user_id + kmean_group_short, agg, lu)
colnames(agg) = c("user_id","kmean_group","n_cycles")
agg_w = reshape(agg, timevar = "kmean_group",idvar = "user_id", direction = "wide")
agg_w[is.na(agg_w)] = 0
users = merge(users, agg_w, all =  TRUE)
rm(agg, agg_w)

# pre-menstrual users
j_prem = which((users$n_cycle_clusters == 1) & (users$n_cycles.PreM == 7) &  (users$n_cycles.M == 0) )[1:10]
j_m = which((users$n_cycle_clusters == 1) & (users$n_cycles.PreM == 0) &  (users$n_cycles.M == 7) )[1:10]
j_mix = which((users$n_cycle_clusters == 2) & (users$n_cycles.PreM == 3) &  (users$n_cycles.M == 4) )[1:10]
j = c(j_prem, j_m, j_mix)
rm(j_prem, j_m, j_mix)
for(i in j){
  cat(users$user_id[i],"\n")
  u = which(d$user_id == users$user_id[i])
  d_subset = d[u,]
  d_subset$cycle_nb_m = factor(d_subset$cycle_nb_m, levels = sort(unique(d_subset$cycle_nb_m), decreasing = TRUE))
  print(ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_nb_m, col = kmean_group))+
          geom_vline(xintercept = 0, col = "gray90", size = 1.5) + 
          geom_point()+ guides(col = FALSE) + ggtitle(paste(users$user_id[i],"-",users$n_cycles.M[i],"-",users$n_cycles.PreM[i]))+
          scale_color_manual(values = cols$clusters)+
          xlab("cycleday from 1st day of menstruation")+ ylab("")+
          scale_x_continuous(breaks = par$x.axis,limits = range(par$x.axis)) )
}
rm(i, j, u)
```


```{r cycle_clustering_enough_logs cycle clusters examples - figure 2B}
users_M = c("01f64acefe89c55cbfe8b0cc931fce9321143210")#,"002974f759b5efd13ecc1740b124ad0f2797ecd5" , "000becd740780336d082f5f982381ac1adc63427"
users_preM = c("09137d795acf3561f37457634c8ab67ab1fef4d9")
users_mix = c("007352b76b20462a8c6d38f17932207308cbc5e6") # "025958bb7d702a73fcbf57606eccd43528229e13","00f25aec0fc7ca2036f65c09087ac94349a9f41c","0034a6c2c80aec6d270bc7f59b79a29935708e6b",
user_ids = c(users_M,users_preM,users_mix)
rm(users_M,users_preM,users_mix)

u = which(d$user_id %in% user_ids)
d_subset = d[u,]
d_subset$cycle_nb_m = factor(d_subset$cycle_nb_m, levels = sort(unique(d_subset$cycle_nb_m), decreasing = TRUE))
d_subset$user_id_f = c("user 1","user 2","user 3")[as.numeric(factor(d_subset$user_id, levels = user_ids))]

g = ggplot(d_subset, aes(x = cycleday_m_D, y = cycle_nb_m, col = kmean_group))+
  facet_grid(user_id_f ~ ., scale = "free_y" )+
  geom_vline(xintercept = 0, col = "gray90", size = 1.5) + 
  geom_point(shape = 17)+ guides(col = FALSE) +
  xlab("cycleday from 1st day of menstruation")+ ylab("cycle #")+
  scale_x_continuous(breaks = par$x.axis,limits = range(par$x.axis)) + 
  scale_color_manual(values = cols$clusters)+
  theme(strip.background = element_rect(fill=rgb(89,155,227,60,max = 255), color = NA))
g

ggsave(filename = paste0(IO$panels, "2B_cycle_clusters_examples.pdf"), plot = g, width = viz$full_width/3, height = viz$full_width/5)

```


### Cycle clusters by age and birth control


```{r cycle_clustering_enough_logs user clusters by age and birth control - cycle clusters Fig 2C }
m = match(d$user_id, users$user_id)
d$age_cat = users$age_cat[m]
d$birth_control = users$birth_control[m]
rm(m)


agg = aggregate(tender_breasts ~ cycleday_m_D + birth_control + age_cat + kmean_group , d, sum)
agg_cycles = aggregate(cycle_id ~ birth_control + age_cat , d, lu)
colnames(agg_cycles)[colnames(agg_cycles) == "cycle_id"] = "n_cycles"
agg = merge(agg, agg_cycles, all = TRUE)
agg$perc_tender_breasts = agg$tender_breasts/agg$n_cycles


g = ggplot(agg[(!agg$age_cat %in% par$age_cat_exclude) & !((agg$age_cat == "(35,40]") & (agg$birth_control == "pill")),], 
           aes(x = cycleday_m_D, y = perc_tender_breasts, col = kmean_group))+
  geom_line()+facet_grid(age_cat ~ birth_control)+
  scale_x_continuous(breaks = par$x.axis)+
  scale_y_continuous(labels = scales::percent)+
  scale_color_manual(values = cols$clusters)+
  xlab("cycleday from 1st day of menstruation") + ylab("n tender breasts / n cycles")+
  guides(col = FALSE)
g

ggsave(filename = paste0(IO$panels, "2C_cycle_clusters_by_age_and_BC.pdf"), plot = g, width = viz$full_width/3*1.2, height = viz$full_width/4.75*1.2)


rm(g, agg, agg2, agg_cycles)

```



### checking the tracking behavior in these 2 clusters


```{r cycle_clustering_enough_logs loading cycle clusters, eval = FALSE}
load("Rdata/cycles_with_clusters.Rdata", verbose = TRUE)
```



```{r cycle_clustering_enough_logs load one days file with tracking behavior}
rm(days)
load(paste0(IO$Rdata,"days/days_1.Rdata"), verbose = TRUE)
head(days)
```



```{r cycle_clustering_enough_logs tracking behavior in the 2 clusters on just one day file}

days$cluster = cycle_clusters$kmean_group[match(days$cycle_id_m, cycle_clusters$cycle_id_m)]
d = days[!is.na(days$cluster),]

breaks$n_logs = c(0:35+0.5, 100)

#h = hist(d$number[(d$cycleday_m == 0)&(d$type == "n_logs")], breaks = breaks$n_logs)
#h = hist(d$number[(d$cycleday_m == -18)&(d$type == "n_logs")], breaks = breaks$n_logs)

agg = aggregate(number ~ cycleday_m_D + cluster, d[d$type == "n_logs",],
                function(x){h = hist(x, breaks = breaks$n_logs, plot = FALSE); h$counts})
agg.tmp = as.data.frame(agg$number)
colnames(agg.tmp) = c(paste0("n_logs_",1:36))

sum_cycleday_cluster = apply(agg.tmp, 1, sum)

n_cycles = aggregate(cycle_id_m ~ cluster, d, lu)
n_cycles = n_cycles[match(agg$cluster, n_cycles$cluster),]

agg = cbind( data.frame(cluster = agg$cluster, 
                        cycleday_m_D = agg$cycleday_m_D, 
                        n_logs_0 = (n_cycles$cycle_id_m - sum_cycleday_cluster)/n_cycles$cycle_id_m), 
             agg.tmp/n_cycles$cycle_id_m)


agg2 = reshape(agg, idvar = c("cluster", "cycleday_m_D"), varying = list(grep("n_logs_",colnames(agg))), direction = "long")
agg2$time = agg2$time - 1
colnames(agg2) = c("cluster","cycleday_m_D", "n_logs","perc")


ggplot(agg2, aes(x = cycleday_m_D, y = n_logs, col = log(perc))) + geom_point() + facet_grid(. ~ cluster)

ggplot(agg2, aes(x = cycleday_m_D, y = n_logs, col = perc)) + geom_point() + facet_grid(. ~ cluster)


ggplot(agg2[agg2$n_logs == 0,], aes(x = cycleday_m_D, y = perc, col = cluster)) + geom_line() 
ggplot(agg2[agg2$n_logs == 0,], aes(x = cycleday_m_D, y = 1-perc, col = cluster)) + geom_line() 


ggplot(agg2[agg2$n_logs %in% 1:3,], aes(x = cycleday_m_D, y = perc, col = cluster)) + geom_line() + facet_grid(n_logs ~. ) 


rm(n_cycles, sum_cycleday_cluster, agg.tmp, agg, agg2, d)


```
